---
title: 'Merging spreadsheets [<i class="fa fa-github" aria-hidden="true"></i>](https://github.com/SteveViss/Talk_bib)'
date: "`r paste0(format(Sys.time(), '%d %B %Y'))`"
author: 'Steve Vissault and Kevin Cazelles'
# lang: french
output:
    ioslides_presentation:
        incremental: false
        widescreen: true
        includes:
          in_header: kcheader.html  
        #logo: ./img/logo.png
---


```{r dsetup, results = "hide", echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.align='center',
  tidy=FALSE,
  comment="R>> ",
  size = 'small'
)
```

## Collecting informations together 



## One spreadsheet

- dealing with `csv` files

```r
read.csv('file.csv')
```

- dealing directly with `.xlsx` files (requires java `sudo R CMD javareconf`):

```r
install.packages('xlsx')
read.xlsx('file.csv')
```


## Multiple spreadsheets

- Use the same headers, put them in a folder then:

```r
fcsv <- list.files('path/to/folder', pattern='^F.*csv$', full.names=TRUE)

lscsv <- list()

for (i in 1:length(fcsv)){
  lscsv <- read.csv(fcsv[[i]])
}
mydat <- do.call('rbind', lscsv)
```

## Multiple spreadsheets


```bash
ssconvert -S fileIn.xlsx foldercsv/fileOut_%s.csv
```

# Google spreadsheet

## Go online

Show how to export them as `.csv` files.



# Alternatives (may be may not be)

## YAML

## JSON

## SQL












## Centered

<div class="centered">
This text is centered.
</div>


## Example with 2 columns

<div class="columns-2">

  - To make it look good
  - I changed the width of the figure  

  ```{r codechunk3, echo=F, fig.width=5.5}
  seqx <- seq(0,10,0.1)
  plot(seqx, cos(seqx), type='h')
  ```

</div>
